---
title: "Make census data tables"
author: "Dave"
format: html
editor: visual
---

## Introduction

### Libraries

```{r libraries}
#| message: false

library(here)
library(dplyr)
library(readxl)
library(fs) # check this actually gets used - I'd like to try it
library(stringr)
library(purrr)
library(readr)
```

## Read data using excel_extract()

Currently we've got a lot of data tables from a previous project. We've also got a function, `excel_extract()` , which is designed to read those tables with some degree of flexibility, because they're not always consistently formatted.

### Load excel_extract()

`excel_extract()` is saved as an RDS in the 'functions' directory.

```{r loadExcelExtract}
excel_extract <- readRDS(here("functions", "excel-extract.RDS"))
```

### Define paths

```{r definePaths}
## taking advantage here's (undocument) list-handling behaviour
census_here <- list("data", "inputs", "census")

nations_path  <- here(census_here, "nations")
dioceses_path <- here(census_here, "dioceses")
coms_path     <- here(census_here, "com")

paths <- list(
  nations  = nations_path,
  dioceses = dioceses_path,
  coms     = coms_path
)
```

### List excel files

```{r listExcels}
excels <- lapply(paths, dir, full.names = TRUE, pattern = "\\.xlsx")
```

### Convert lists to vectors and name

Currently, `excels` is a 3-length list of vectors. The simplest way to run `excel_extract()` on each file will be to convert the vectors within each list to a set of named vectors

```{r convertListsToVectors}
listify_files <- function(x){
  list_names <-
    str_extract(x, "(?<=/)[^/]+(?=$)") |> #final file name
    str_remove(".xlsx") |>
    str_replace_all("&", "and") |> 
    str_remove_all("[:punct:]") |> 
    str_squish() |> 
    str_replace_all("[:space:]", "-")
  
  out <- as.list(x)
  names(out) <- list_names
  
  return(out)
}

excels <- lapply(excels, listify_files)

show <- function(x) print(str_c(class(x), length(x), sep = "; "))
walk(excels,show)
```

### Read excel files

It's *probably* best to do this one directory at at a time, since there are only 3 directories and each one might want slightly different arguments. NB this code block takes maybe like 15/20 minutes to run.

```{r readExcels}
#| cache: true
census_tables <- lapply(excels, lapply, excel_extract,
                        tabs = c(
                          1, "Ethnicity", "Religion", "Health", "Household type"
                        ),
                        numberise = TRUE)
```

## Write CSVs

### Create directories

```{r createDirs}
dir_create(unlist(paths), "csvs")

walk2(
  census_tables,
  paths,
  function(table_list, path){
    path <- path(path, "csvs")
    
    location_names <- as.list(names(table_list))
    
    walk2(table_list, location_names,
          
      function(location, location_name){
        
        sub_dir <- dir_create(path, location_name)
        
        stat_names <- as.list(names(location))
        
        walk2(location, stat_names,
              
              function(stat, stats_name){
                write_csv(
                  stat,
                  str_c(
                    sub_dir,
                    str_c(
                      location_name, "-", stats_name, ".csv"
                      ),
                    sep = "/"
                  ))})
      })
  })
    
```
